Learning features and abstract actions for computing generalized plans

### 2018-09-05 doc/paper.tex
draft progress

In the experiments we will   consider also  a smaller theory  that avoids the consideration of all pairs
of expanded states or all pairs of transitions in $\cal S$]. For this, some expanded states $s$ and transitions $(s,s')$ in
tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by computing some plans for some instances $P$ in $\Q$, 
and by adding the resulting state transitions to ${\cal S}$ at the same time that the first
arguments $s$ and $(s,s')$ in the  $D_1$ and $D_2$ atoms are restricted to be marked.
If the set of marked states and transitions is much smaller than the set of unmarked ones, 
then the resulting formula,  $T_G({\cal S},{\cal F})$ will be much smaller than $T({\cal S},{\cal F})$.
The goal marking scheme that involves the theory $T_G({\cal S},{\cal F})$ in place of $T({\cal S},{\cal F})$
amounts to enforcing soundness on  the sample set ${\cal S}$ but not completeness, yet this does
not affect the resulting  formal guarantees that depend on soundness only.

@@

The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=10$, unless no solution is found
to the theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
% We have considered strategies to increase the bound adaptively
The resulting theory $T({\cal S},{\cal F})$ is solved with the Open-WBO MAXSAT solver \cite{martins2014open}.

@@

\begin{table*}[t]
\centering
% \addtolength{\tabcolsep}{-1pt}
{
% \scriptsize
% \footnotesize
\small
\def\arraystretch{1}%  1 is the default

@@

\rowspacing \hline \rowspacing
\multirow{1}{*}{$\Q_{on\text{-}2\text{-}tow}$}   
& 3 &        420  & 8 & 657 & -K & -K & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18

@@

\paragraph{Clearing a block.}
The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
in the Blocks world. 
We take the standard encoding with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.

The abstract action model features one action \emph{pick-above-$x$} to pick the topmost block above $x$,
and one action \emph{place-aside} to put it away, both described in terms of features $H$, $X$ and $n(x)$, 
that denote, respectively, whether any block is being held, whether block $x$ is being held,
and the number of blocks above $x$.\footnote{
Action and features names are assigned manually by us to increase readability, but correspond faithfully
to the learnt entities.
}
%
The model is provably correct, but not complete in general, as it lacks e.g.
any action to pick block $x$.
However, it is sufficient to obtain a policy with a single loop that will 
alternatively \emph{pick-above-$x$} and \emph{place-aside} until block $x$ is clear,
which provably \emph{achieves the goal from any configuration of blocks, for any number of blocks}.

@@

\paragraph{Stacking two blocks from different towers.}
The problem $\Q_{on\text{-}2\text{-}tow}$ is aims at achieving a goal of the form $on(x,y)$
in the Blocks world for any two blocks $x$ that are initially not in the same tower.
We take the same encoding as above.

\paragraph{Gripper.}
The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two arms needs to 
move a number of blocks between two different rooms, each arm being able to carry only one ball at a time.
We use the standard encoding with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
plus type unary predicates $room$, $ball$, and $gripper$.

\paragraph{Collecting rewards in a maze.}
The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
The encoding has primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.

** Mention: symbold DRL paper by Shanahan in

>>>>

In the experiments we will consider also a smaller theory  that avoids the consideration of all pairs
of expanded states or all pairs of transitions in $\cal S$. For this, some expanded states $s$ and transitions $(s,s')$ in
tuples $(s,s',P)$ are marked as \emph{goal relevant}. This is achieved by computing some plans for some instances $P$ in $\Q$, 
and by adding the resulting state transitions to ${\cal S}$ at the same time that the first
arguments $s$ and $(s,s')$ in the  $D_1$ and $D_2$ atoms are restricted to be marked.
If the set of marked states and transitions is much smaller than the set of unmarked ones, 
then the resulting formula,  $T_G({\cal S},{\cal F})$ will be much smaller than $T({\cal S},{\cal F})$.
The goal marking scheme that involves the theory $T_G({\cal S},{\cal F})$ in place of $T({\cal S},{\cal F})$
amounts to enforcing soundness on  the sample set ${\cal S}$ but not completeness, yet this does
not affect the resulting  formal guarantees that depend on soundness only.

@@

The feature complexity bound $k$ in the feature pool $\mathcal{F}$ is set to a value of $k=8$.
In our experiments, we increased this value to $k=16$ if no model was found
for the logical theories $T$ and $T_G$, and although it would be more effective
to devise an adaptive strategy to do so incrementally or through binary search,
that is beyond the scope of this work.
% theory $T({\cal S},{\cal F})$, in which case it is increased to $k=20$. 
% We have considered strategies to increase the bound adaptively
The resulting theories are then solved with the Open-WBO MAXSAT solver \cite{martins2014open}.

@@

\begin{table*}[t]
\centering
% \addtolength{\tabcolsep}{-1pt}
{
% \def\arraystretch{1}%  1 is the default
\small

@@

\rowspacing \hline \rowspacing
\multirow{1}{*}{$\Q_{on\text{-}2\text{-}tow}$}   
& 3 &        420  & 8 & 657 & 128K & 25.8M & 18.3K &     3.3M    &  0.02&  5 & 7   &  - & - & - & - \\% Total complexity of F: 18
% Max-sat problem has 128217 variables and 25825428 unique clauses (with repetitions: 25825428). Top weight is 4577

@@

\paragraph{Clearing a block.}
The generalized problem $\Q_{clear}$ aims at achieving the goal $clear(x)$ for any block $x$
in the Blocks world. 
We take the standard encoding with primitive predicates $on(a, b)$, $clear(b)$, $ontable(b)$, $holding(b)$ and $handempty$.
%
With one single instance $\Q_{clear}$ with 5 blocks, we are able to learn
an abstract action model which features two actions:

- \emph{pick-above-$x$}: \tuple{\neg X, \neg H, n(x) > 0; n(x) \downarrow},

- \emph{place-aside}: \tuple{\neg X, H; \neg H},

\noindent both described in terms of features $H: \abs{holding}$, $X: \abs{holding \sqcap \set{x}}$ and $n(x)\equiv \abs{\exists on^* . \set{x}}$, 
that denote, respectively, whether any block is being held, whether block $x$ is being held,
and the number of blocks above $x$.\footnote{
Action and features names are assigned manually by us to increase readability, but correspond faithfully
to the learnt entities.
}
%
The model is provably correct, but not complete in general, as it lacks e.g.
any action to pick block $x$.
However, it is sufficient to obtain a policy with a single loop that will 
alternatively \emph{pick-above-$x$} and \emph{place-aside} until block $x$ is clear,
which provably \emph{achieves the goal from any configuration of blocks, for any number of blocks}.

@@

\paragraph{Stacking two blocks from different towers.}
The problem $\Q_{on\text{-}2\text{-}tow}$ is aims at achieving a goal of the form $on(x,y)$
in the Blocks world for any two blocks $x$ and $y$ that are initially not in the same tower.
We take the same encoding as above. Based on 3 different instances with up to 14
blocks, we are able to learn an abstract action model with 5 features and 7 actions.
The set of features is made up of  $X$, $n(x)$ and $n(y)$, with the same syntax and semantics as given before,
as well as $E: handempty$
and $XY: [(\exists on^{-1} . \set{x})\sqcap \set{b}]$,
the second of which precisely denotes that block $x$ is on block $y$.
The set of actions includes actions to pick up the topmost block above $x$, to pick up the topmost block above $y$,
to put a block being held \emph{aside}, i.e. not above $x$ nor above $y$, to pick $x$, 
to place $x$ on $y$ and to place $x$ on the table. This last action will indeed be necessary in case block $x$ is initially being held
but block $y$ is not clear.

DISCUSS POLICY WHEN AVAILABLE. / 
DISCUSS WHY WOULDN'T WORK WHEN IN SAME TOWER.

\paragraph{Gripper.}
The problem $\Q_{gripper}$ is based on the classical Gripper problem, where a robot with two grippers needs to 
move a number of blocks from one room into another \emph{target} room $x$, which we take to be a parameter of the problem,
each gripper of the robot being able to carry only one ball at a time.
We use the standard encoding with primitive predicates $at\text{-}robby(l)$, $at\text{-}ball(b, l)$, $free(g)$, $carry(b, g)$ that denote,
respectively, whether the robot (the ball) is at location $l$, whether gripper $g$ is free, and whether gripper $g$ carries ball $b$,
plus unary type predicates $room$, $ball$, and $gripper$.
%
We feed our pipeline with 2 training instances of the problem with only 2 rooms and at most 4 balls.
The resulting abstract action model is based on the following features:

- $C: \exists carry . \top$ (number of balls carried by the robot),

- $X: at\_robby \sqcap \set{x}$ (whether robot in target room),

- $G: \abs{free}$ (number of empty grippers), and

- $B: \exists at . \neg \set{x}$ (number of balls not in target room),

\noindent and contains the following actions :

- \emph{leave-ball(s)-at-$x$}: \tuple{C>0, X;  C\downarrow, G \uparrow}

- \emph{move-to-$x$-half-loaded}: \tuple{\neg X, B=0, C>0, G>0;   X}

- \emph{move-to-$x$-fully-loaded}: \tuple{\neg X, C>0, G=0;   X}

- \emph{pick-ball-not-in-$x$}: \tuple{\neg X, B > 0, G > 0;   B \downarrow, G \downarrow, C\uparrow}

- \emph{leave-$x$}: \tuple{X, C=0, G > 0;   \neg X}

The model is not complete, as it e.g. features no action to pick a ball from room $x$,
but is goal-oriented as a result of using the SAT theory $T_G$, in that it features all of the actions
necessary to achieve the goal.
This can be seen e.g. in the fact that the action to move \emph{move-to-$x$-half-loaded}
can only be performed when no more balls remain in the other room.
Additionally, the model is \emph{provably correct for any gripper instance with an arbitrary number of balls, 
rooms and grippers}. CHECK !!

The policy that results ... (!!??)

% Action 1:
%         PRE: ncarried > 0, robot-at-B
%         EFFS: DEC ncarried, INC nfree-grippers
% 
% Action 2:
%         PRE: NOT robot-at-B, nballs-A = 0, ncarried > 0, nfree-grippers > 0
%         EFFS: robot-at-B
% 
% Action 3:
%         PRE: NOT robot-at-B, ncarried > 0, nfree-grippers = 0
%         EFFS: robot-at-B
% 
% Action 4:
%         PRE: NOT robot-at-B, nballs-A > 0, nfree-grippers > 0
%         EFFS: DEC nballs-A, DEC nfree-grippers, INC ncarried
% 
% Action 5:
%         PRE: ncarried = 0, nfree-grippers > 0, robot-at-B
%         EFFS: NOT robot-at-B

\paragraph{Collecting rewards in a maze.}
The objective in the problem $\Q_{reward}$ is to move in an $n \times m$ grid to pick up certain
rewards that are spread out in certain grid cells, while avoiding other cells which are considered as \emph{blocked}.
The encoding has primitive predicates $reward(c)$, $at(c)$, $blocked(c)$ and $adjacent(c_1, c_2)$ that denote
respectively the position of the rewards, of the agent, of the blocked cells, and the grid topology.
We use two single instances of the problem with size $5 \times 5$ and $4 \times 4$ and different distributions of blocks and rewards.
We learn a set with 2 features:

- $R: \abs{reward}$ (number of remaining rewards), and

- $D: dist(at, adjacent:(\neg blocked), reward)$ (distance from current cell to closest cell with reward, traversing adjacent, unblocked cells only).

\noindent The action model has also 2 actions:

- \emph{pick-reward}: \tuple{D=0, R > 0;  R \downarrow, D \uparrow}

- \emph{move-towards-closest-reward}: \tuple{R>0, D>0; D \downarrow}

The model is not complete ... 
...

The policy ... 
... EXPAND A BIT - TOO TIRED NOW.
Discuss suboptimality of the policy.
Perhaps discuss D=$\infty$ when no reward?

% Action 1 [PICK-REWARD]:
% 	PRE: min-dist[at, Restrict(adjacent,Not(blocked)), reward] = 0, num-rewards > 0
% 	EFFS: DEC num-rewards, INC min-dist[at, Restrict(adjacent,Not(blocked)), reward]
% 
% Action 2 [MOVE-TO-CLOSEST-REWARD]:
% 	PRE: min-dist[at, Restrict(adjacent,Not(blocked)), reward] > 0, num-rewards > 0
% 	EFFS: DEC min-dist[at, Restrict(adjacent,Not(blocked)), reward]

\paragraph{Summary.}
XXX Perhaps mention general aspects of the results here, if space allows, e.g.:
\begin{itemize}
 \item Running times of the SAT solver are negligible.
 \item Minimization of precondition DNFs
 \item Theory T
 \item Sample sparsity - we use very few instances, rather few transitions.
 \item ** Mention: symbold DRL paper by Shanahan in 
\end{itemize}
